---
title: "IST687 - Music Classification Project"
author: "Team 2 - Sebastian Castro, John Fields, Courtney Smith, Jeremy Wallner"
date: "5/13/2019"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

The purpose of this project is to analyze the Million Song Database to predict "Hot" artists and songs based on the attributes such as familiarity, artist location, loudness, terms used, etc.  The analysis was done using R software on a 10,000 track subset of the data and our model was able to predict "Hot" songs with ~80% accuracy. 

## Table of Contents

Executive Summary
Data
Analysis
Conclusion
Final proofing

## Introduction

## Related Work

Thierry Bertin-Mahieux, Daniel P.W. Ellis, Brian Whitman, and Paul Lamere. 
The Million Song Dataset. In Proceedings of the 12th International Society
for Music Information Retrieval Conference (ISMIR 2011), 2011.

## Dataset
```{r import data}

#New code from Courtney to change from 3 to 5 categories of artist hotness
#music <- read.csv("~/Intro data science/Music project/newmusic.csv")
music <- read.csv("/Users/johnfields/Library/Mobile Documents/com~apple~CloudDocs/Syracuse/IST687/Project/Data Files/think.cs.vt.edu/music.csv")
str(music)
colnames(music)[1] <- "artist.hotttnesss"

#Plot of the variables
library(ggplot2)
library(reshape2)
ggplot(data = melt(music), mapping = aes(x = value)) + geom_histogram(bins = 20) + facet_wrap(~variable, scales = 'free_x')


#New code from Jeremy importing of song list
#newmusic <- read.csv("~/Intro data science/Music project/newmusic3.csv")
newmusic <- read.csv("/Users/johnfields/Library/Mobile Documents/com~apple~CloudDocs/Syracuse/IST687/Project/Data Files/think.cs.vt.edu/newmusic3.csv")

#head(newmusic)
newmusic2 <- newmusic
newmusic3 <- newmusic2[-c(1:2,4:9,11,13:14,19:20,23,36)]

newmusic3 <- na.omit(newmusic3)
cmbomusic <- newmusic3


##Artist Hotness Histogram
library(ggplot2)
ggplot(music, aes(x=artist.hotttnesss))  + geom_histogram(color="black", fill="steelblue", alpha=0.5) + xlab("Artist Hotness") + ggtitle("Histogram: Artist Hotness")
##Function to create descriptive statistics for artist hotness
descriptive_stats <- function(vector) { library(moments) 
  result <- c(Mean=mean(vector),
              Median=median(vector),
              Min = min(vector),
              Max = max(vector),
              SD = sd(vector),
              Quantile = quantile(vector, probs = c(0.25,.50,0.75, 0.95)),
              Skewness = skewness(vector) )
  print(result) 
}
descriptive_stats(music$artist.hotttnesss)
##Methodology for assigning artist hotness levels - uses quantiles from descriptitive_statistics function
#95% Quantile: 0.6011861 - Hot
#75% Quantile: 0.453858  - Warm
#50% Quantile: 0.3807423 - Tepid
#25% Quantile: 0.3252656 - Cool
##Code for assigning labels based on above quantiles
music$artist.hotness.label <- ifelse(music$artist.hotttnesss >=0.6011861, "Hot",
                                     ifelse(music$artist.hotttnesss >=0.453858 & music$artist.hotttnesss <0.6011861, "Warm",
                                            ifelse(music$artist.hotttnesss >=0.3807423 & music$artist.hotttnesss <0.453858, "Tepid",
                                                   ifelse(music$artist.hotttnesss >=0.3252656 & music$artist.hotttnesss <0.3807423, "Cool",
                                                          ifelse(music$artist.hotttnesss < 0.3252656, "Frigid","Else")))))
unique(music$artist.hotness.label)
#End of new code from Courtney
#Prior to importing, a new column artist.hotttnesss.label was adding with 
#Hot(>.4590), Warm(<.4590 and >.3357), Cold(<.3357).  Four rows with blanks in 
#famiiarity were also deleted.
```


```{r cleaning}
music <- na.omit(music)
#Copy original data to a new dataframe music1 and exclude unneeded data
music <- music[-c(2:5,7,16,19,21:25,30,34)]
music$artist.hotness.label <- as.factor(music$artist.hotness.label) 
str(music)

##SONG HOTNESS HISTOGRAM From Jeremy
cmbomusic[cmbomusic==0]<- NA
#cmbomusic2 <- cmbomusic[-c(5,6)]
cmbomusic3 <- na.omit(cmbomusic)
cmbomusic3$song.hotttnesss.label <- ifelse( cmbomusic3$song.hotttnesss >=0.6011861, "Hot",ifelse(cmbomusic3$song.hotttnesss >= 0.453858 & cmbomusic3$song.hotttnesss <0.6011861, "Warm", ifelse(cmbomusic3$song.hotttnesss >=0.3807423 & cmbomusic3$song.hotttnesss < 0.453858, "Tepid",ifelse(cmbomusic3$song.hotttnesss >=0.3252656 & cmbomusic3$song.hotttnesss <0.3807423, "Cool",ifelse(cmbomusic3$song.hotttnesss < 0.3252656, "Frigid","Else")))))
unique(cmbomusic3$song.hotttnesss.label)
cmbomusic3 <- cmbomusic3[-c(2:3,12)]
ggplot(cmbomusic3, aes(x=song.hotttnesss))  + geom_histogram(color="black", fill="steelblue", alpha=0.5) + xlab("Song Hotness") + ggtitle("Histogram: Song Hotness")
##Function to create descriptive statistics for song  hotness
descriptive_stats2 <- function(vector) { library(moments) 
  result <- c(Mean=mean(vector),
              Median=median(vector),
              Min = min(vector),
              Max = max(vector),
              SD = sd(vector),
              Quantile = quantile(vector, probs = c(0.25,.50,0.75, 0.95)),
              Skewness = skewness(vector) )
  print(result) 
}
descriptive_stats2(cmbomusic3$song.hotttnesss)

cmbomusic3$song.hotness.label <- ifelse( cmbomusic3$song.hotttnesss >=0.64787976, "Hot",ifelse(cmbomusic3$song.hotttnesss >= 0.43437984  & cmbomusic3$song.hotttnesss <0.64787976, "Warm", ifelse( cmbomusic3$song.hotttnesss <0.43437984, "Cold","Else")))
unique(cmbomusic3$song.hotness.label)


cmbomusic3$song.hotttnesss.label <- as.factor(cmbomusic3$song.hotttnesss.label) 
str(cmbomusic3)

cmbomusic3$song.hotttnesss.label <- ifelse( cmbomusic3$song.hotttnesss >=0.6011861, "Hot",ifelse(cmbomusic3$song.hotttnesss >= 0.453858 & cmbomusic3$song.hotttnesss <0.6011861, "Warm", ifelse(cmbomusic3$song.hotttnesss >=0.3807423 & cmbomusic3$song.hotttnesss < 0.453858, "Tepid",ifelse(cmbomusic3$song.hotttnesss >=0.3252656 & cmbomusic3$song.hotttnesss <0.3807423, "Cool",ifelse(cmbomusic3$song.hotttnesss < 0.3252656, "Frigid","Else")))))
unique(cmbomusic3$song.hotttnesss.label)

cmbomusic3$song.hotttnesss.label <- as.factor(cmbomusic3$song.hotttnesss.label) 
str(cmbomusic3)

cmbomusic3$song.hotttnesss.label <- ifelse( cmbomusic3$song.hotttnesss >=0.6011861, "Hot",ifelse(cmbomusic3$song.hotttnesss >= 0.453858 & cmbomusic3$song.hotttnesss <0.6011861, "Warm", ifelse(cmbomusic3$song.hotttnesss >=0.3807423 & cmbomusic3$song.hotttnesss < 0.453858, "Tepid",ifelse(cmbomusic3$song.hotttnesss >=0.3252656 & cmbomusic3$song.hotttnesss <0.3807423, "Cool",ifelse(cmbomusic3$song.hotttnesss < 0.3252656, "Frigid","Else")))))
unique(cmbomusic3$song.hotttnesss.label)

str(cmbomusic3)

cmbomusic3$song.hotttnesss.label <- as.factor(cmbomusic3$song.hotttnesss.label) 
cmbomusic3$song.hotttnesss.label <- as.factor(cmbomusic3$song.hotttnesss.label) 
str(cmbomusic3)
 #View the number of Cold/Warm/Hot labels 
 table(cmbomusic3$song.hotttnesss.label)

 cmbomusic3$song.hotness.label <- ifelse( cmbomusic3$song.hotttnesss >=0.64787976, "Hot",ifelse(cmbomusic3$song.hotttnesss >= 0.43437984  & cmbomusic3$song.hotttnesss <0.64787976, "Warm", ifelse( cmbomusic3$song.hotttnesss <0.43437984, "Cold","Else")))
unique(cmbomusic3$song.hotness.label)

```
## Features

```{r features}
#View the number of Cold/Warm/Hot labels 
table(music$artist.hotttnesss.label)
#View the number of Frigid/Cool/Tepid/Warm/Hot labels 
table(music$artist.hotness.label)
#Plot artists latitude and longitude
plot(music$latitude,music$longitude)

cmbomusic3$song.hotness.label <- as.factor(cmbomusic3$song.hotness.label) 
cmbomusic3$song.hotness.label <- as.factor(cmbomusic3$song.hotness.label) 
str(cmbomusic3)
#View the number of Cold/Warm/Hot labels 
table(cmbomusic3$song.hotness.label)
#View the number of Frigid/Cool/Tepid/Warm/Hot labels 
 table(cmbomusic3$song.hotttnesss.label)
 #Plot artists latitude and longitude
 plot(cmbomusic3$latitude,cmbomusic3$longitude)
#Plot artist hotttnesss
#hist(music$artist.hotttnesss,breaks=20)
#hist(music$artist.hotness,breaks=20)
```
#Create a map of the world
mapWorld <- borders("world", colour="gray50", fill="white")

```{r world map}
#New code from John for creating a map of the world showing latitude/longitude and artist hotness
#Code based on info from https://rpubs.com/spoonerf/global_map
library(dplyr)
loc<-data.frame(music$longitude,music$latitude,music$artist.hotttnesss)
loc<-unique(loc)
colnames(loc)<-c("longitude", "latitude","artist hotness")
loc_df<-data.frame(loc)
library(maps)
library(mapdata)
library(ggplot2)
ahworld <- ggplot(data=loc_df, aes(longitude, latitude, group=NULL,fill=NULL,size=artist.hotness))+#, fill=hole)) + 
  borders(fill="light grey",colour="light grey")+
  geom_point(color="black",alpha=I(7/10))+
  scale_size(range=c(1,7), guide = "legend",labs(size="Artist Hotness"))+
  coord_equal()+  ggtitle("Artist Location by Hotness")
ahworld
```
```{r}
#New code from John for creating a map of the world showing latitude/longitude and artist hotness
#Code based on info from https://rpubs.com/spoonerf/global_map
library(dplyr)
songlc<-data.frame(cmbomusic3$longitude,cmbomusic3$latitude,cmbomusic3$song.hotttnesss)
songlc<-unique(songlc)
colnames(songlc)<-c("longitude", "latitude","song hotness")
songlc_df<-data.frame(songlc)
library(maps)
library(mapdata)
library(ggplot2)
songlc_dfwrld <- ggplot(data=songlc_df, aes(longitude, latitude, group=NULL,fill=NULL,size=song.hotness))+#, fill=hole)) + 
  borders(fill="light grey",colour="light grey")+
  geom_point(color="black",alpha=I(7/10))+
  scale_size(range=c(1,7), guide = "legend",labs(size="Song Hotness"))+
  coord_equal()+  ggtitle("Song Location by Hotness")
songlc_dfwrld
```

## Methods - Linear Regression

```{r methods 1}
# code from Juan
lm(formula = music$artist.hotttnesss ~ music$year + music$bars_confidence +
    + music$tempo + music$duration + music$start_of_fade_out +
   music$tatums_start + music$familiarity + music$latitude +
   music$tempo + music$longitude + music$beats_start + music$beats_confidence +
   music$end_of_fade_in)
#removed music$bars_start which was causing an error

```
  
## Methods - Random Forest

```{r methods 2}
#Do analysis to determine hot/warm/cold artists based on hotttnesss
#The ramdom forest analysis is from a training video by Bharatendra Rai 
#at https://www.youtube.com/watch?v=dJclNIN-TPo
#Data Partition - ind = independent samples
#The code below runs in console but not R Markdown
set.seed(123)
ind<- sample(2,nrow(music), replace=TRUE,prob=c(0.7,0.3))
train <- music[ind==1,]
test <- music[ind==2,]
#Run randomForest on 3 levels
library(randomForest)
set.seed(222)
rf <- randomForest(music[,-21:-22],music[,21])
print(rf)
attributes(rf)
rf$confusion
#Run randomForest on 5 levels
library(randomForest)
set.seed(222)
rf2 <- randomForest(music[,-21:-22],music[,22])
print(rf2)
attributes(rf2)
rf2$confusion

#Error rate of Random Forest
plot(rf)
plot(rf2)
#The error rate is not improving after ~100 trees

# Variable Importance
# Familiarity is much more important than the other variables.
varImpPlot(rf,
           sort=T,
           n.var=10,
           main="Top 10 - Variable Importance")
importance(rf)
varUsed(rf)
varImpPlot(rf2,
           sort=T,
           n.var=10,
           main="Top 10 - Variable Importance")
importance(rf2)
varUsed(rf2)

cmbomusic4 <- na.omit(cmbomusic3)
cmbomusic5 <- cmbomusic4[-c(1,3,7:9,13,17,20)]
str(cmbomusic5)
cmbomusic5$song.hotness.label <- as.factor(cmbomusic4$song.hotness.label) 
rf3 <- randomForest(cmbomusic5[,-12:-13],cmbomusic5[,13])
rf3
print(rf3)
attributes(rf3)
rf3$confusion
plot(rf3)

varImpPlot(rf3,
           sort=T,
           n.var=10,
           main="Top 10 - Variable Importance")
```
## Results

## Conclusion

## Appendices