---
title: "IST687 - Music Classification Project"
author: "Team 2 - Sebastian Castro, John Fields, Courtney Smith, Jeremy Wallner"
date: "4/18/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary

## Table of Contents

## Introduction

## Related Work

Thierry Bertin-Mahieux, Daniel P.W. Ellis, Brian Whitman, and Paul Lamere. 
The Million Song Dataset. In Proceedings of the 12th International Society
for Music Information Retrieval Conference (ISMIR 2011), 2011.

## Dataset
```{r import data}
#New code from Courtney to change from 3 to 5 categories of artist hotness
music <- read.csv("/Users/johnfields/Library/Mobile Documents/com~apple~CloudDocs/Syracuse/IST687/Project/Music Project/think.cs.vt.edu/newmusic.csv")

str(music)

##Artist Hotness Histogram
library(ggplot2)
ggplot(music, aes(x=artist.hotttnesss))  + geom_histogram(color="black", fill="steelblue", alpha=0.5) + xlab("Artist Hotness") + ggtitle("Histogram: Artist Hotness")

##Function to create descriptive statistics for artist hotness
descriptive_stats <- function(vector) { library(moments) 
  result <- c(Mean=mean(vector),
              Median=median(vector),
              Min = min(vector),
              Max = max(vector),
              SD = sd(vector),
              Quantile = quantile(vector, probs = c(0.25,.50,0.75, 0.95)),
              Skewness = skewness(vector) )
  print(result) 
}

descriptive_stats(music$artist.hotttnesss)

##Methodology for assigning artist hotness levels - uses quantiles from descriptitive_statistics function
#95% Quantile: 0.6011861 - Hot
#75% Quantile: 0.453858  - Warm
#50% Quantile: 0.3807423 - Tepid
#25% Quantile: 0.3252656 - Cool

##Code for assigning labels based on above quantiles
music$artist.hotness.label <- ifelse(music$artist.hotttnesss >=0.6011861, "Hot",
                                     ifelse(music$artist.hotttnesss >=0.453858 & music$artist.hotttnesss <0.6011861, "Warm",
                                            ifelse(music$artist.hotttnesss >=0.3807423 & music$artist.hotttnesss <0.453858, "Tepid",
                                                   ifelse(music$artist.hotttnesss >=0.3252656 & music$artist.hotttnesss <0.3807423, "Cool",
                                                          ifelse(music$artist.hotttnesss < 0.3252656, "Frigid","Else")))))

unique(music$artist.hotness.label)
#End of new code from Courtney

#Prior to importing, a new column artist.hotttnesss.label was adding with 
#Hot(>.4590), Warm(<.4590 and >.3357), Cold(<.3357).  Four rows with blanks in 
#famiiarity were also deleted.

remove_na <- function(music,n=0)
  {
  music <- music[rowSums(is.na(music)) <= n,]
  music <- as.data.frame.matrix(music)
  return (music)
  } 

#Copy original data to a new dataframe music1 and exclude unneeded data
music <- music[-c(1:5,7,16,19,21:25,30,34)]
music$artist.hotness.label <- as.factor(music$artist.hotness.label) 
str(music)
```
## Features

```{r features}
#View the number of Cold/Warm/Hot labels 
table(music$artist.hotttnesss.label)

#View the number of Frigid/Cool/Tepid/Warm/Hot labels 
table(music$artist.hotness.label)

#Plot artists latitude and longitude
plot(music$latitude,music$longitude)

#Plot artist hotttnesss
#hist(music$artist.hotttnesss,breaks=20)
#hist(music$artist.hotness,breaks=20)
```
#THIS IS INCOMPLETE CODE FOR PLOTTING ADDITIONAL DATA...
#Create a map of the world
mapWorld <- borders("world", colour="gray50", fill="white")

#Code from https://rpubs.com/spoonerf/global_map
#Need to figure out what to put in locs
locs<-read.csv("my_locations.csv")
locs<-
sp_dups<-data.frame(ddply(locs,.(Longitude,Latitude),nrow))
sp_dups$loc_id<-1:length(sp_dups$Longitude)
sp_dups_df<-merge(sp_dups, locs, by=c("Longitude","Latitude"))

loc<-data.frame(sp_dups_df$Longitude,sp_dups_df$Latitude,sp_dups_df$V1)
loc<-unique(loc)
colnames(loc)<-c("Longitude", "Latitude", "V1")

coordinates(loc)<-c("Longitude","Latitude")
proj4string(loc) <- CRS("+proj=longlat")

loc_df<-data.frame(loc)

theme_opts <- list(theme(panel.grid.minor = element_blank(),
                         panel.grid.major = element_blank(),
                         panel.background = element_blank(),
                         plot.background = element_rect(fill="white"),
                         panel.border = element_blank(),
                         axis.line = element_blank(),
                         axis.text.x = element_blank(),
                         axis.text.y = element_blank(),
                         axis.ticks = element_blank(),
                         axis.title.x = element_blank(),
                         axis.title.y = element_blank(),
                         plot.title = element_text(size=22)))

library(maps)
library(mapdata)

ggplot(data=loc_df, aes(Longitude, Latitude, group=NULL,fill=NULL,size=V1))+#, fill=hole)) + 
  borders(fill="light grey",colour="light grey")+
  geom_point(color="black",alpha=I(7/10))+
  scale_size(range=c(1,7), guide = "legend",labs(size="No. of Populations"))+
  coord_equal()+
  theme_opts
  
## Methods
```{r methods}
#Do analysis to determine hot/warm/cold artists based on hotttnesss

#The ramdom forest analysis is from a training video by Bharatendra Rai 
#at https://www.youtube.com/watch?v=dJclNIN-TPo
#Data Partition - ind = independent samples
#The code below runs in console but not R Markdown
set.seed(123)
ind<- sample(2,nrow(music), replace=TRUE,prob=c(0.7,0.3))
train <- music[ind==1,]
test <- music[ind==2,]

#Run randomForest on 3 levels
library(randomForest)
set.seed(222)
rf <- randomForest(music[,-21:-22],music[,21])
print(rf)
attributes(rf)
rf$confusion

#Run randomForest on 5 levels
library(randomForest)
set.seed(222)
rf2 <- randomForest(music[,-21:-22],music[,22])
print(rf2)
attributes(rf2)
rf2$confusion

#Run randomForest again with tune mtry data from below
#Need HELP to fix the next line of code so it works...
#rf <- randomForest(artist.hotness.label ~.,data=music1,ntree=200,mtry=8,
#importance=TRUE,proximity=TRUE)

#Prediction & Confusion Matrix - train data
#library(caret)
#p1<-predict(rf,train)
# For some reason this is returning an error buit p2 below is working
#confusionMatrix(p1,train)

#Predition & Confusion Matrix - test data
#p2<-predict(rf,test)
#confusionMatrix(p2,test$artist.hotness.label)

#Error rate of Random Forest
plot(rf)
plot(rf2)
#The error rate is not improving after ~100 trees

#Tune mtry
#t <- tuneRF(train[,-21],train[,21],
#           stepFactor=.5,
#            plot=TRUE,
#            ntreeTry=200,
#            trace=TRUE,
#            improve=0.05)

#No. of nodes for the trees
hist(treesize(rf),
    main="Number of Nodes for the Trees",
    col="green")

hist(treesize(rf2),
    main="Number of Nodes for the Trees",
    col="green")

# Variable Importance
# Familiarity is much more important than the other variables.  Should it be removed and run again?
varImpPlot(rf,
           sort=T,
           n.var=10,
           main="Top 10 - Variable Importance")
importance(rf)
varUsed(rf)

varImpPlot(rf2,
           sort=T,
           n.var=10,
           main="Top 10 - Variable Importance")
importance(rf2)
varUsed(rf2)

#Mulit-dimenstional Scaling Plot
#The code below causes R to lock up...
#MDSplot(rf,train$artist.hotttnesss.label)

```
## Results

## Conclusion

## Appendices